---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# actordupes

<!-- badges: start -->
<!-- badges: end -->

The goal of actordupes is to provide a set of synchronised functions to filter duplicates from datasets that already contain actors who are already categorised with regards to the questions whether they are the author of an article (journalist), non-human entities (misclassification) or passive (no direct/indirect quote). It was developed to preprocess actor datasets that had been coded by a local Large Language Model for further manual coding without manually exclude duplicates in that stage of data collection.
**Important note**: As it is common for journalistic texts to provide full names (sur- and last name) and only last names for most actors, as well as exclusively surnames for actors who, for example, wish to remain anonymous, the detection of duplicates is mainly based on last names and/or the last given name. So if there is just a surname available it will compare the surname of the person with last names or people whose full name is included. In some cases, this can lead to wrong classifications of duplicates. The functions in the package are not designed to be applied to datasets containing mostly/exclusively surnames.


## Installation

You can install the development version of actordupes from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("tablue99/actordupes")
```

## Example

The basic structure of datasets set out to be cleaned with the functions in this package needs to contain at least four variables. Firstly, a distinct ID for each actor (entity_id), their name (entity), an ID of the article in which they were identified (document_id) and a variable containing the preliminary decision (of the LLM) on whether the actor is relevant (relevant).

```{r dataframe example, echo = FALSE}
entity_id <- c(1:5)
entity <- c("Philipp Müller", "Miley Cyrus", "Cyrus", "Wagner", "Richard Wagner")
document_id <- c(1, 1, 1, 1, 1)
relevant <- c(TRUE, TRUE, TRUE, FALSE, TRUE)

actors_dataframe <- data.frame(entity_id, entity, document_id, relevant)
actors_dataframe
```

To receive a clean version of the actor dataset with a correct indication of active, thus relevant actors, you have to follow these xx steps shown in this code example:

### 1. Load the package
```{r step 1}
library(actordupes)
```

### 2. Generate a adjusted copy of your dataframe with "prepare_duplicate_dataframe()"

This will generate a version of your actors dataframe that contains the lowercase names of the entities (entity_low) as well as two columns containing the first and last names of the actors.

```{r step 2}
prepared_actors_dataframe <- prepare_duplicate_dataframe(actors_dataframe)

prepared_actors_dataframe
```

### 3. Mark actors that appear more than once within a document (duplicates) with "find_duplicates()"

This will add a column called "duplicate" to your initial dataframe. The values are "TRUE" for actors who already appeared within the article and thus are duplicates and "NA" for actors who either appear only once or, in case of multiple appearances, at their first appearance. Note that you can set the threshold of similarity (common strings) individually. Values from 0 (no similarity) to 1 (full similarity) are accepted. Typically, values of .75 or higher are set to detect duplicates in a reliable manner.
Always remember to assign the applied function to your initial or a new dataframe to see the changes as there will be no other output in the console.

```{r step 3}
actors_dataframe <- find_duplicates(actors_dataframe, prepared_actors_dataframe, 0.8)

actors_dataframe
```

### 4. Change the coding in relevant subsequent to the detection of duplicates with "mark_relevant_actors()"

If you not only wish to detect duplicates but also use this information to identify relevant actors, it is not enough to simply filter for actors who are not duplicates (is.na(duplicate)) as not all of them are necessarily quoted at their first mention which will lead to a loss of actors who are quoted at their second, third or later appearance for the first time and therefore, are relevant. To check whether an actor that is not quoted at their first appearance is quoted later throughout the article, the package contains the function "mark_relevant_actors()" which only changes the value of relevance for duplicates from TRUE to FALSE if the actor has also been relevant at the time of his first mention. To do so, again the similarity of names, predominantly last names, is used to group the actors within an article.

```{r step 4}
actors_dataframe <- mark_relevant_actors(actors_dataframe, prepared_actors_dataframe, 0.8)

actors_dataframe
```

Note that the value in relevant for "3 = Cyrus" has been changed while "5 = Richard Wagner" remains the same.
**ATTENTION**: The "mark_relevant_actors()" function will irreversibly change the values in the "relevant" column of the initial dataframe. If you wish to keep it, assign the resulting dataframe to a copy of the initial dataframe. If you wish to have both categorisations of relevant in one dataframe, you can subsequently join them via their corresponding entity IDs (e.g., by using "left_join()").